---
title: "Overview"
output: rmarkdown::html_vignette
bibliography: "bibliography.bib"
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 4,
  fig.height = 4,
  fig.align = "center"
)
```

```{r setup}
library(rasenna)
data(belcho)
```

## Resample telemetry data

Many analyses using telemetry data assume that the fixes have been sampled at the same temporal resolution, i.e. at a constant time interval.
The `resample()` function re-sample telemetry data to force this assumption.
Re-sampling is done by removing fixes that are closer in time than an user-specified time interval (see _freq_ and _low_ in `?resample()`).
When a fix is removed, the time interval between fixes is re-calculated.
Additionally, fixes that are apart more than a given time interval (see _high_ in `?resample()`) are kept, but they are assigned different _ID_s to indicate they belong to different tracks and should not be analyzed as part of the same consecutive trajectory.

```{r resample}
belcho[["timestamp"]] <- strptime(
  belcho[["timestamp"]], 
  format = "%Y-%m-%d %H:%M:%S", 
  tz = "UTC"
)
res <- resample(belcho, ts_col = "timestamp", freq = 3, units = "hours")
summary(res[["dt"]])
```


## Test for serial correlation of direction movement

Telemetry data that is at high temporal resolution may violate the assumption of independence of contiguous fixes.
This can be detected by assessing the correlation between contiguous fixes [@moorcroft2013mechanistic].

```{r direction}
cor.test.direction(res[res[["ID"]] == 1, c("x", "y")], method = "pearson")
```
# Localizing tendency model

The localizing tendency models assumes that direction angles are distributed following a Von Mises distribution:
\[
  \text{VM}(\phi, \hat\phi, \kappa)= \frac{exp(\kappa \cdot cos(\phi - \hat\phi))}{2 \pi I_0(\kappa)}
\]
and relocation steps follow a negative exponential distribution:
\[
  f_\tau(\rho, \lambda) = \lambda \cdot exp(-\lambda\cdot \rho)
\]

The parameters $\kappa$ and $\lambda = 1/\bar{\rho}$ can be estimated directly from the data [@moorcroft2013mechanistic, p. 143-144].

```{r capra}
data(capra)
capra[["timestamp"]] <- strptime(
  capra[["timestamp"]], 
  format = "%Y-%m-%d %H:%M:%S", 
  tz = "UTC"
)
res <- resample(capra, ts_col = "timestamp", freq = 6, units = "hours")
res <- res[res$ID == 309, ] #keep only track 309
xy <- res[, c("x", "y")]
cor.test.direction(xy) #test serial correlation
```

```{r ltm_angle}
k <- ltm_kappa(direction(xy), relative_direction(xy))
hist(
  direction(xy), 
  probability = TRUE, 
  breaks = 10, 
  xlab = "direction (rad)",
  main = ""
)
lines(
  seq(-pi, pi, by = .01), 
  von_mises(
    seq(-pi, pi, by = .01),
    mean(direction(xy), na.rm = TRUE),
    kappa = k
  ),
  lw = 2, col = "darkred"
)
```

```{r ltm_steps}
rho <- ltm_rho(xy)
dx <- xy[-1, 1] - xy[-nrow(xy), 1]
dy <- xy[-1, 2] - xy[-nrow(xy), 2]
d <- sqrt(dx^2 + dy^2)
hist(
  d, 
  probability = TRUE, 
  breaks = 10,
  xlab = "Step length (m)",
  main = ""
)
lines(
  seq(0, 1200, length.out = 5e2),
  exponential(seq(0, 1200, length.out = 5e2), 1 / rho),
  lw = 2, col = "darkred"
)
```

Assumptions of modeling angles and step-length using the Von Mises and negative exponential distributions should always be checked.
The data from _Capra ibex_ used above, for example, do not follow a Von Mises distribution and using the localizing tendency model is questionable.

# References
